---
title: "R Notebook"
output:
  html_document:
    df_print: paged
runtime: shiny
---

Let's illustrate a pretty useful method to model periodic series: generating artifitial fourier series to capture the cycles and sub-cycles on a periodic time series
```{r}
source('/Users/ssobrinou/IE/Advanced/2019_Advanced/Regression/code/load_libraries.R')

data("AirPassengers")
plot(AirPassengers); grid()

y<-copy(AirPassengers)

df<-data.table(date=seq(as.Date('1949-01-01'), as.Date('1960-12-01'), by='1 month'),y=y)

train_df<-df[date<'1960-01-01']
test_df<-df[date>='1960-01-01']

train_y<-ts(train_df$y,freq=12,start=c(1949,1))

```

The stl decomposition method from the forecast package show's us a de-composition of our series onto trend, seasonality (cycle) and irregular
```{r}
library(forecast)

str(train_y)
frequency(train_y)

# stl function for time series de-composiong

decomp<-stl(train_y, s.window = 12)
plot(decomp); grid()
```

The fourier function generates k-pairs of fourier transformations based on the dependent variable time frecuency
```{r}
# creating fourier series of different amplitude

forecast::fourier(train_y,K=4)[1:10,]

df_fourier<-data.table(train_y=train_y,
               fourier(train_y,K=2))
colnames(df_fourier)<-gsub('-','_',colnames(df_fourier))
head(df_fourier)
```


Actualy, fourier series are just sin and cosine functions for 1, 2...k amplitude of an angular function
```{r}

t=seq(1,12)
N=12

c1<-cos(2*pi*t/N)
s1<-sin(2*pi*t/N)

c2<-cos(2*2*pi*t/N)
s2<-sin(2*2*pi*t/N)

identical(round(c1,4), round(df_fourier$C1_12[1:12],4))
identical(round(s1,4), round(df_fourier$S1_12[1:12],4))
identical(round(c2,4), round(df_fourier$C2_12[1:12],4))
identical(round(s2,4), round(df_fourier$S2_12[1:12],4))

```


```{r}
par(mfrow=c(1,2))
plot(c1, type='b', main='1st & 2nd Sin Wave'); grid()
lines(c2, type='b', col='blue')
plot(s1, type='b', main='1st & 2nd Sin Wave'); grid()
lines(s2, type='b',  col='blue')
par(mfrow=c(1,1))
```



Now we fit a linear regression using the fourier terms to
capture the seasonallity of the series on an automatic manner
```{r}
df_model<-data.table(date=seq(as.Date('1949-01-01'), as.Date('1960-12-01'), by='1 month'),
                           y=as.numeric(y),
                           fourier(y,K=4))
      df_model[, tend:=seq(1,nrow(df_model))]
      
      df_model_train<-df_model[date<'1960-01-01']
      df_model_test<-df_model[date>='1960-01-01']
      
      # estimate model
      lm_fourier<-lm(log(y)~., data=df_model_train[, -'date'])
      summary(lm_fourier)
      
plot(as.numeric(train_y), type='l'); grid()
lines(exp(fitted(lm_fourier)), col='red')
```


The number of fourier terms shjould be optimised on the bias-variance tradeoff:
```{r, echo=FALSE}
library(shiny)
shinyApp(
  
  ui = fluidPage(
    sliderInput(label="Number of Fourier Series",
                inputId =  "k",
                min=1,
                max=12,
                value=1,
                step=1),
    plotOutput("fitplot")
  ),
  
  server = function(input, output) {
    output$fitplot = renderPlot({
      
      # calculate fourier series varying with k input (and a fixed linear trend)
      df_model<-data.table(date=seq(as.Date('1949-01-01'), as.Date('1960-12-01'), by='1 month'),
                           y=as.numeric(y),
                           fourier(y,K=input$k))
      df_model[, tend:=seq(1,nrow(df_model))]
      
      df_model_train<-df_model[date<'1960-01-01']
      df_model_test<-df_model[date>='1960-01-01']
      
      # estimate model
      lm_fourier<-lm(log(y)~., data=df_model_train)
      
      # plot fit
      plot(as.numeric(train_y), type='l'); grid()
      lines(exp(fitted(lm_fourier)), col='red')
    })
  },
  
  options = list(height = 500)
)
```

Predicting with our model: 

```{r}
test_df[, pred_fourier:=round(exp(predict(lm_fourier, newdata=df_model_test)))]
test_df

ggplot(melt(test_df, id.vars = 'date'), aes(x=date, y=value, colour=variable))+geom_point()+geom_line()

```


Another option...go into Arima!!!
```{r}
arima_model<-auto.arima(train_y,lambda = 'auto')
summary(arima_model)

test_df[, pred_arima:=round(as.data.frame(forecast(arima_model, h=12))[,1])]

ggplot(melt(test_df, id.vars = 'date'), aes(x=date, y=value, colour=variable))+geom_point()+geom_line()

```


Also interesting trying with seasonal dummy variables: 
```{r}
df[, month:=factor(lubridate::month(date))]
str(df)

data_months<-caret::dummyVars(formula= ~., data = df, fullRank=T,sep = "_")
data_months<-data.table(predict(data_months, newdata = df))


data_months[, date:=seq(as.Date('1949-01-01'), as.Date('1960-12-01'), by='1 month')]


str(data_months)

data_months[, tend:=seq(1, nrow(data_months))]

train_df_months<-data_months[date<'1960-01-01']
test_df_months<-data_months[date>='1960-01-01']

```


```{r}
lm_months<-lm(log(y)~., data=train_df_months)
summary(lm_months)

test_df[, pred_months:=round(exp(predict(lm_months, newdata = test_df_months)))]

ggplot(melt(test_df, id.vars = 'date'), aes(x=date, y=value, colour=variable))+geom_point()+geom_line()

```


