---
title: "Interactive Visualization - Advanced R"
author: "Sirio Sobrino"
date: "20/4/2019"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
    incremental: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('/Users/ssobrinou/IE/Advanced/2019_Advanced/Regression/code/load_libraries.R')
library(htmlwidgets, quietly = T)
library(highcharter, quietly = T)
library(plotly, quietly = T)
library(ggiraph, quietly = T)
library(leaflet, quietly=T)

target="price"

```

## R Presentation with ioslides

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Some powerfull libraries for Interactive Visualization {.incremental}

- plotly
- ggiraph
- dygraphs
- leaflet
- highcharter
- visNetwork
- networkD3



## Reading the data {.smaller .flexbox .vcenter} 
```{r reading_data} 
path='/Users/ssobrinou/IE/Advanced/2019_Advanced/Datasets/Regression/'
file='Automobile_2.csv'
df<-fread(paste(path,file, sep='/'),
          stringsAsFactors = T)
df<-df[complete.cases(df)]
head(df,5)
```

## plotly 

- You can use pure plotly which is pretty flexible and powerful.
- Limited features compared to ggplot.

```{r plotly}

plot_ly(df, x=~hp, y=~get(target), type='scatter', mode='markers')

```


## plotly::ggplotly

- Plotly has a wraper function to turn any ggplot chart interactive.
- For complicated charts it's not easy to customize and doesn't deal very well with shiny in some cases. 
- Still a very recommended function for standard charts.

```{r ggplot,fig.height=3, fig.width=4.5}

ggplotly(ggplot(df, aes(x=hp, y=get(target)))+
           geom_hex()+
           geom_smooth(method = 'loess',formula= 'y ~ x')+
           ylab(target))


```

## ggiraph

- A library designed to make ggplot interactive.
- Doesn't support all ggplot features but the vast majority.
- More stable and easier to tune than plotly.

```{r ggiraph}
p<-ggplot(df, aes(x=hp, y=get(target),colour=make, tooltip=paste(sep = "<br/>",
                                                           make,
                                                           paste('hp:',hp),
                                                           paste0(target,': ',hp))))+
  ylab(target)+
  geom_point_interactive()


```

---
```{r plot_ggiraph}
ggiraph::ggiraph(code={print(p)}, width = 0.5, width_svg = 7)
```

## highcharter 

```{r highcharter}
hchart(df, type="scatter", hcaes(x = hp, y = get(target), group = drive))
```

--- 

```{r rCharts}
# hchart(density(df$hp), type="area")
hchart(df, type="column", hcaes(x=hp,y=get(target),group = drive))

```